---
title: "Assignment 01"
author: "Mengyao"
date: "2020/9/5"
output: 
  html_document:
    theme: lumen
    toc: true
    toc_float: true
---
# 1. Load tidycensus, tidyverse, dplyr, httr, knitr
```{r message=FALSE}
library(tidycensus)
library(tidyverse)
library(dplyr)
library(httr)
library(knitr)
```

# 2. Load household-level data from the 1-year American Community Survey Sample from 2018
```{r}
household_variable_2018 <- pums_variables %>%
  distinct(year, survey, var_code, var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```

# 3. Fix connection problem
## I am constantly getting an error message "Failed to connect to api.census.gov". I have searched online for different methods. This chunk solves my problem.
```{r results='hide'}
 set_config(
     use_proxy(url="127.0.0.1", port=7890)
 )
 httr::GET("https://www.census.gov/")
```

# 4. Choose 5 variables:
## I want to do a household-level analysis. After looking through the variables available in household_variable_2018, I 've chosen the following five variables:

     - Food stamp (FS) (categorical)
     - Laptop (LAPTOP) (categorical)
     - Household income (HINCP) (continuous)
     - Number of own children (NOC) (continuous)
     - Property Value (VALP) (continuous)
     
## I want my sample to include all households in New York State.
```{r results= "hide"}
household_data <- get_pums(variables = c("FS", "LAPTOP", "HINCP", "NOC", "VALP"),
                           state = "NY",
                           year = 2018,
                           survey = "acs1",
                           recode = TRUE) 
```

# 5. Select and present columns from household_data dataframe
## This dataset includes 196,967 households. First 10 rows are presented here as a formatted table.
```{r}
hh_data <- select(household_data, "FS", "LAPTOP", "HINCP", "NOC", "VALP")
knitr::kable(hh_data[1:10,])
```

