---
title: "Assignment 01"
author: "Mengyao"
date: "2020/9/5"
output: 
  html_document:
    theme: lumen
    toc: true
    toc_float: true
---
# Loading packages
```{r message=FALSE}
library(tidycensus)
library(tidyverse)
library(dplyr)
library(httr)
library(knitr)
```

# Loading household-level data from the 1-year American Community Survey Sample from 2018

I want to do a household-level analysis. 
```{r}
household_variable_2018 <- pums_variables %>%
  distinct(year, survey, var_code, var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```

# Fixing connection problem

I am constantly getting an error message "Failed to connect to api.census.gov". I have searched online for different methods. This chunk solves my problem.
```{r results='hide'}
 set_config(
     use_proxy(url="127.0.0.1", port=7890)
 )
 httr::GET("https://www.census.gov/")
```

# Choosing 6 variables

After looking through the variables available in household_variable_2018, I have chosen the following six variables:

     - When structure first built (YBL)(categorical)
     - Household language (HHL) (categorical)
     - Access to the internet (ACCESS)(categorical)
     - Household income (HINCP) (continuous)
     - First mortgage payment (MRGP) (continuous)
     - Property Value (VALP) (continuous)
     
I want my sample to include all households in New York State. 
Through those data, I would like to explore the relationship between household income and property value, structure age, mortgage payment, and household language. The following questions can find an answer through this analysis: "Do richer families spend more on housing?", "Do they tend to live in newer structure?", "Do all households use English?", etc.   
```{r results= "hide"}
household_data <- get_pums(variables = c("YBL", "HHL", "ACCESS", "HINCP", "MRGP", "VALP"),
                           state = "NY",
                           year = 2018,
                           survey = "acs1",
                           recode = TRUE) 
```

# Selecting and presenting columns from household_data dataframe

This dataset includes 79,116 households. The first ten rows are presented here as a formatted table.
```{r}
hh_data <- household_data %>%
  filter(YBL != "bb",
         MRGP != 0,
         ACCESS != "b") %>%
  mutate(intel_access = case_when(ACCESS == "1"~"TRUE",
                                  ACCESS == "2"~"TRUE",
                                  ACCESS == "3"~"FALSE")) %>%
  mutate(household_language = case_when(HHL == "1"~"English only",
                                        HHL == "2"~"Spanish",
                                        HHL == "3"~"Other Indo-European languages",
                                        HHL == "4"~"Asian and Pacific Island languages",
                                        HHL == "5"~"Other language")) %>%
  mutate(bd_year = case_when(YBL == "01"~"Really old",
                             YBL == "02"~"Really old",
                             YBL == "03"~"Really old",
                             YBL == "04"~"Old",
                             YBL == "05"~"Old",
                             YBL == "06"~"Old",
                             YBL == "07"~"Old",
                             YBL == "08"~"New",
                             YBL == "09"~"New",
                             YBL == "10"~"New",
                             YBL == "11"~"New",
                             YBL == "12"~"New",
                             YBL == "13"~"New",
                             YBL == "14"~"New",
                             YBL == "15"~"Really new",
                             YBL == "16"~"Really new",
                             YBL == "17"~"Really new",
                             YBL == "18"~"Really new",
                             YBL == "19"~"Really new",
                             YBL == "20"~"Really new",
                             YBL == "21"~"Really new",
  )) %>%
  select(bd_year, household_language, intel_access, "HINCP", "MRGP", "VALP")

hh_data
```

# Saving data
```{r}
write_csv(hh_data, "household data_02.csv")
```

