---
title: "Assignment 01"
author: "Mengyao"
date: "2020/9/5"
output: html_document
---
# 1. Load tidycensus, tidyverse, dplyr
```{r eval=FALSE}
library(tidycensus)
library(tidyverse)
library(dplyr)
```

# 2. Load household-level data from the 1-year American Community Survey Sample from 2018
```{r}
household_variable_2018 <- pums_variables %>%
  distinct(year, survey, var_code, var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```

# 3. Choose 5 variables:
     - Food stamp (FS) (categorical)
     - Laptop (LAPTOP) (categorical)
     - Household income (HINCP) (continuous)
     - Number of own children (NOC) (continuous)
     - Property Value (VALP) (continuous)
```{r}
household_data <- get_pums(variables = c("FS", "LAPTOP", "HINCP", "NOC", "VALP"),
                           state = "NY",
                           year = 2018,
                           survey = "acs1",
                           recode = TRUE) 
```
# 4. Select and present columns from household_data dataframe
```{r}
hh_data <- select(household_data, "FS", "LAPTOP", "HINCP", "NOC", "VALP")
hh_data
```

# 5. Saving data
```{r}
write_csv(hh_data, "household data.csv")
```

