---
title: "Assignment 03"
author: "Mengyao"
date: "2020/9/18"
output: 
  html_document:
    theme: lumen
    toc: true
    toc_float: true
---

# Loading packages
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
```

# Loading data
Loading the household-level dataset from week 2. 
This is a household-level analysis. The data are collected from the 1-year American Community Survey Sample from 2018 and covered all households in New York that spent money on the first mortgage payment.

the following five variables are included:
     - When structure first built (bd_year)(categorical)
     - Household language (household_language) (categorical)
     - Household income (HINCP) (continuous)
     - First mortgage payment (MRGP) (continuous)
     - Property Value (VALP) (continuous)
     
I filter those buildings whose construction time is unclear, and rewrite it as csv file for future use. 
```{r, message=FALSE}
household_data_as2 <- read_csv("household data_02.csv")
household_data_as3 <- household_data_as2 %>%
  filter(bd_year != "NA")
write_csv(household_data_as3, "household data_03.csv")
```

# Relationship between two continuous variables

1. Relationship between household income (HINCP) and first mortgage payment (MRGP)
```{r}
cor_income_mortgage <- cor.test(household_data_as3$HINCP, household_data_as3$MRGP)
cor_income_mortgage
```

visualizing the relationship
```{r}
ggplot(household_data_as3,
       aes(x = HINCP, y = MRGP)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_light() +
  scale_x_continuous(name = "Household Income (dollars)") +
  scale_y_continuous(name = "First Mortgage Payment (dollars)") +
  annotate("text", x = 0, y = 6000,
           label = paste("Correlation (95-percent confidence): ", prettyNum(cor_income_mortgage$conf.int[1], digits =2), "to", prettyNum(cor_income_mortgage$conf.int[2])),
           hjust = 0)
```

The correlation between household income and first mortgage payment is positive and quite strong. With higher income, families tended to spend more on first mortgage payment. 

2. Relationship between household income (HINCP) and property value (VALP)
```{r}
cor_income_property <- cor.test(household_data_as3$HINCP, household_data_as3$VALP)
cor_income_property
```

visualizing the relationship
```{r}
options(scipen = 999)
ggplot(household_data_as3,
       aes(x = HINCP, y = VALP)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_light() +
  scale_x_continuous(name = "Household Income (dollars)") +
  scale_y_continuous(name = "Property Value (dollars)") +
  annotate("text", x = 0, y = 4500000,
           label = paste("Correlation (95-percent confidence): ", prettyNum(cor_income_property$conf.int[1], digits =2), "to", prettyNum(cor_income_property$conf.int[2])),
           hjust = 0)
```

The correlation between household income and property value is positive and quite strong. Richer families tended to live in more expensive houses. 

3. Relationship between first mortgage payment (MRGP) and property value (VALP)
```{r}
cor_mortgage_property <- cor.test(household_data_as3$MRGP, household_data_as3$VALP)
cor_mortgage_property
```

visualizing the relationship
```{r}
options(scipen = 999)
ggplot(household_data_as3,
       aes(x = MRGP, y = VALP)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_light() +
  scale_x_continuous(name = "First Mortgage Payment (dollars)") +
  scale_y_continuous(name = "Property Value (dollars)") +
  annotate("text", x = 0, y = 4500000,
           label = paste("Correlation (95-percent confidence): ", prettyNum(cor_mortgage_property$conf.int[1], digits =2), "to", prettyNum(cor_mortgage_property$conf.int[2])),
           hjust = 0)
```

The correlation between first mortgage payment and property value is positive and quite strong. More expensive houses often require higher first mortgage payment. Yet I am expecting the correlation to be even stronger. I am guessing that maybe when houses are not that expensive (less than 1000000 dollars), the price have little influence on the first mortgage payment.

# Relationship between a continuous variable and a categorical variable (three or more levels)

I would like to thank Sara for clarification of the Tukey HSD test. 

1. Relationship between household income (HINCP) and when structure first built (bd_year)
```{r}
anova_income_bdage <- aov(HINCP ~ bd_year, data = household_data_as3)
summary(anova_income_bdage)
```

The p-value is much lower than 0.05, so at a 95-percent confidence interval, there is a significant association between the household income and structure age.

Taking a Tukey's Honestly Significant Difference test to see the magnitude of the differences among the difference categories.
```{r}
differences_income_bdage <- TukeyHSD(anova_income_bdage)
as_tibble(cbind(pair = row.names(differences_income_bdage$bd_year),
                differences_income_bdage$bd_year))
```

Only for the pairing of "Really new" and "New", the p-value is higher than 0.05, which means that there is little association with the household income. 

Visualizing those differences with a box plot
```{r}
ggplot(household_data_as3,
       aes(x = bd_year, y = HINCP)) +
  geom_boxplot() +
  theme_light() +
#  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "When Structure First Built") +
  scale_y_continuous(name = "Household Income",
                     breaks = seq(0, 1750000, by = 250000),
                     labels = paste("$", 
                                    prettyNum(seq(0, 1750000, by = 250000), 
                                              big.mark = ",")))
```

2. Relationship between household income (HINCP) and household language (household_language)
```{r}
anova_income_language <- aov(HINCP ~ household_language, data = household_data_as3)
summary(anova_income_language)
```

The p-value is much lower than 0.05, so at a 95-percent confidence interval, there is a significant association between the household income and household language.

Taking a Tukey's Honestly Significant Difference test to see the magnitude of the differences among the difference categories.

```{r}
differences_income_language <- TukeyHSD(anova_income_language)
as_tibble(cbind(pair = row.names(differences_income_language$household_language),
                differences_income_language$household_language))
```

For the pairing of "other language" and "Other Indo-European languages", "Spanish" and "English only", the p-value is higher than 0.05, which means that there is little association with the household income. 

Visualizing those differences with a violin plot
```{r}
ggplot(household_data_as3,
       aes(x = household_language, y = HINCP)) +
  geom_violin() +
  theme_light() +
  scale_x_discrete(name = "Household Language",
                   labels = c("Asian and Pacific \nIsland languages", "English only", "Other Indo-European \nlanguages", "Other language", "Spanish")) +
  scale_y_continuous(name = "Household Income",
                     breaks = seq(0, 1750000, by = 250000),
                     labels = paste("$", 
                                    prettyNum(seq(0, 1750000, by = 250000), 
                                              big.mark = ",")))
```

3. Relationship between first mortgage payment (MRGP) and when structure first built (bd_year)
```{r}
anova_mortgage_bdage <- aov(MRGP~bd_year, data = household_data_as3)
summary(anova_mortgage_bdage)
```
The p-value is much lower than 0.05, so at a 95-percent confidence interval, there is a significant association between the first mortgage payment and structure age.

Taking a Tukey's Honestly Significant Difference test to see the magnitude of the differences among the difference categories.

```{r}
differences_mortgage_bdage <- TukeyHSD(anova_income_bdage)
as_tibble(cbind(pair = row.names(differences_mortgage_bdage$bd_year),
                differences_mortgage_bdage$bd_year))
```

Only for the pairing of "Really new" and "New", the p-value is higher than 0.05, which means that there is little association with the first mortgage payment. 

Visualizing those differences with a box plot
```{r}
ggplot(household_data_as3,
       aes(x = bd_year, y = MRGP)) +
  geom_boxplot() +
  theme_light() +
  scale_x_discrete(name = "When Structure First Built") +
  scale_y_continuous(name = "First Mortgage Payment",
                     breaks = seq(0, 7500, by = 1250),
                     labels = paste("$", 
                                    prettyNum(seq(0, 7500, by = 1250), 
                                              big.mark = ",")))
```

4. Relationship between first mortgage payment (MRGP) and household language (household_language)
```{r}
anova_mortgage_language <- aov(MRGP ~ household_language, data = household_data_as3)
summary(anova_mortgage_language)
```

The p-value is much lower than 0.05, so at a 95-percent confidence interval, there is a significant association between the first mortgage payment and household language.

Taking a Tukey's Honestly Significant Difference test to see the magnitude of the differences among the difference categories.

```{r}
differences_mortgage_language <- TukeyHSD(anova_mortgage_language)
as_tibble(cbind(pair = row.names(differences_mortgage_language$household_language),
                differences_mortgage_language$household_language))
```

Only for the pairing of "other language" and "Asian and Pacific Island languages", the p-value is higher than 0.05, which means that there is little association with first mortgage payment. 

Visualizing those differences with a violin plot
```{r}
ggplot(household_data_as3,
       aes(x = household_language, y = MRGP)) +
  geom_violin() +
  theme_light() +
  scale_x_discrete(name = "Household Language",
                   labels = c("Asian and Pacific \nIsland languages", "English only", "Other Indo-European \nlanguages", "Other language", "Spanish")) +
  scale_y_continuous(name = "First Mortgage Payment",
                     breaks = seq(0, 7500, by = 1250),
                     labels = paste("$", 
                                    prettyNum(seq(0, 7500, by = 1250), 
                                              big.mark = ",")))
```

5. Relationship between property value (VALP) and when structure first built (bd_year)
```{r}
anova_property_bdage <- aov(VALP~bd_year, data = household_data_as3)
summary(anova_property_bdage)
```
The p-value is much lower than 0.05, so at a 95-percent confidence interval, there is a significant association between the property value and structure age.

Taking a Tukey's Honestly Significant Difference test to see the magnitude of the differences among the difference categories.

```{r}
differences_property_bdage <- TukeyHSD(anova_property_bdage)
as_tibble(cbind(pair = row.names(differences_property_bdage$bd_year),
                differences_property_bdage$bd_year))
```

Visualizing those differences with a box plot
```{r}
ggplot(household_data_as3,
       aes(x = bd_year, y = VALP)) +
  geom_boxplot() +
  theme_light() +
  scale_x_discrete(name = "When Structure First Built") +
  scale_y_continuous(name = "Property Value",
                     breaks = seq(0, 6000000, by = 1000000),
                     labels = paste("$", 
                                    prettyNum(seq(0, 6000000, by = 1000000), 
                                              big.mark = ",")))
```

6. Relationship between property value (VALP) and household language (household_language)
```{r}
anova_property_language <- aov(VALP ~ household_language, data = household_data_as3)
summary(anova_property_language)
```

The p-value is much lower than 0.05, so at a 95-percent confidence interval, there is a significant association between the property value and household language.

Taking a Tukey's Honestly Significant Difference test to see the magnitude of the differences among the difference categories.

```{r}
differences_property_language <- TukeyHSD(anova_property_language)
as_tibble(cbind(pair = row.names(differences_property_language$household_language),
                differences_property_language$household_language))
```

Only for the pairing of "other language" and "Asian and Pacific Island languages", the p-value is higher than 0.05, which means that there is little association with the property value. 

Visualizing those differences with a violin plot
```{r}
ggplot(household_data_as3,
       aes(x = household_language, y = VALP)) +
  geom_violin() +
  theme_light() +
  scale_x_discrete(name = "Household Language",
                   labels = c("Asian and Pacific \nIsland languages", "English only", "Other Indo-European \nlanguages", "Other language", "Spanish")) +
  scale_y_continuous(name = "Property Value",
                     breaks = seq(0, 6000000, by = 1000000),
                     labels = paste("$", 
                                    prettyNum(seq(0, 6000000, by = 1000000), 
                                              big.mark = ",")))
```

# Relationship between two categorical variables

Relationship between when structure first built (bd_year) and household language (household_language)
```{r}
chi_bdage_language <- chisq.test(household_data_as3$bd_year, household_data_as3$household_language)
chi_bdage_language
```
The p-value is much less than 0.05, so the relationship is significant at a 95-percent confidence level. 

Here are the values I observe for combination of variables:
```{r}
chi_bdage_language$observed
```

Here are the values I would expect if the two variables were not related:
```{r}
chi_bdage_language$expected
```
